Internal Gravity Wave Breaking in White Dwarf Binaries

### 2019-08-19 writeups/Su_IGW_break.tex
stash

    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\left\langle#1\right\rangle}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator*{\med}{med}

@@

Compact white dwarf (WD) binary systems, with orbital periods in the range of
minutes to hours, are important for a range of astrophysical problems. They are
the most important sources of gravitational waves (GWs) for the Laser
Interferometric Space Antenna (LISA) \citep{lisa}. They are also thought to
produce interesting optical transients such as underluminous
supernovae \citep{underlum}, Ca-rich fast transients \citep{carich}, and tidal
novae \citep{tidal_novae}. Most importantly, they have been proposed as the likely
progenitors of type Ia supernovae (e.g. \citep{Ia0,webbink} or more
recently \citep{Ia1,Ia2}). While presently only a few tens of compact WD binaries
are known \citep{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
expand the catalog to a few hundreds \citep{lsst_wd} (results based on
\emph{Gaia}'s second data release have already begun to
appear \citep{gaiaDD,gaiaDD2}), and the Large Synoptic Survey Telescope (LSST,
first light scheduled for 2020) will likely detect a few thousand
more \citep{lsst_wd}. These observations will significantly advance the
understanding of WD binaries and their evolution. My proposed theoretical and
computational research is well-timed to take advantage of these new advances.

@@

The Euler equations for an incompressible, barotropic fluid in a uniform
gravitational field are
\begin{subequations}\label{se:nl_orig}
    \begin{align}
        \bm{\nabla} \cdot \bm{u} &= 0,\label{eq:nl_incomp}\\
        \md{\rho}{t} &= 0 ,\label{eq:nl_density}\\
        \md{\bm{u}}{t} + \frac{\bm{\nabla}P}{\rho} + g\bm{z} &=
            0\label{eq:nl_mom}.
    \end{align}
\end{subequations}
$\md{}{t} = \pd{}{t} + \p{\bm{u} \cdot \bm{\nabla}}$ is the Lagrangian or
material derivative, and $\bm{u}, \rho, P$ denote the velocity field, density
and pressure respectively. We denote $-g\bm{z}$ constant gravitational
acceleration. Note that at hydrostatic equilibrium $\pd{}{t} = 0$ we have
$\bm{\nabla}\overline{P} = -\overline{\rho} g\bm{z}$ and so $\overline{P} =
\overline{\rho} gH$. A vertically-stratified shear flow
$\overline{u}_x(z)\bm{x}$ is permitted at hydrostatic equilibrium, but we will
assume no background flow, so $\bm{u} = \bm{u}'$. Physically, this assumption
corresponds to a non-rotating star, or going to the corotating frame of a
rigidly rotating star.

In practice, it is convenient to introduce coordinate $\Upsilon = \ln
\frac{\rho}{\overline{\rho}}$ (e.g.\ see \citep{lecoanet_anel}). This both
identically enforces $\rho > 0$ and avoids numerical issues in the
$\frac{\bm{\nabla}P}{\rho}$ term if $\rho$ is small. We also define reduced
pressure $\varpi = \frac{P}{\rho}$. Then, we may rewrite the second two
equations of \autoref{se:nl_orig} as
\begin{subequations}\label{se:nl_upsilon}
    \begin{align}
        \md{\Upsilon}{t} + u'_z \pd{\ln \overline{\rho}}{z} &= 0
            ,\label{eq:nl_up_density} \\
        \md{\bm{u}'}{t} + \bm{\nabla}\varpi
            + \varpi\bm{\nabla}\Upsilon + g\bm{z} &= 0.
    \end{align}
\end{subequations}
Note that in the new coordinates, hydrostatic equilibrium corresponds to
$\Upsilon = 0, \overline{\varpi} = gH$.

@@

In the small perturbation limit, where flow velocities are small compared to the
characteristic space and time scales $\pd{}{t} \gg \bm{u}' \cdot
\bm{\nabla}$, we may linearize \autoref{se:nl_upsilon}. The solution in this
linear regime up to undetermined amplitude $A$ is given by
\citep{drazin,sutherland0}:
\begin{equation}
    u'_z\p{x, z, t} = Ae^{z/2H}\cos\p{\p{k_{x}x + k_{z}z - \omega t}},
        \label{eq:k0z_sign}
\end{equation}
where $\omega$ satisfies dispersion relation
\begin{equation}
    \omega^2 = \frac{N^2k_{x}^2}{k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}}.
        \label{eq:disp_rel}
\end{equation}
We take
\begin{equation}
    N^2 \equiv g^2\p{\rd{\rho}{P} - \frac{1}{c_s^2}} = \frac{g}{H},
\end{equation}
the \emph{Brunt-V\"ais\"al\"a frequency} to be constant in the incompressible
limit $c_s \to \infty$ where $c_s$ is the sound speed. Other dynamical
quantities are simply related to $u'_z$ (see e.g.\ \cite{sutherland0}).

In the weak stratification limit $\abs{k_{z}H} \gg 1$, the solution exhibits the
following characteristics:
\begin{itemize}
    \item The amplitude of the wave grows with $z$ as $e^{z/2H}$. Thus, the
        linear approximation is always violated for sufficiently large $z$.

    \item The phase and group velocities are respectively:
        \begin{align}
            \bm{c}_{ph} &=
                \p{k_{x}\bm{x} + k_{z}\bm{z}}\frac{\omega}
                {k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}},\\
            \bm{c}_{g} &= N\frac{\p{k_{z}^2 + \frac{1}{4H^2}}\bm{x}
                - \p{k_{x}k_{z}\bm{z}}}
                {\p{k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}}^{3/2}}.\label{eq:vg}
        \end{align}
        We note $\bm{c}_{ph} \cdot \bm{c}_g =
        \mathcal{O}\p{\frac{1}{\p{k_{z}H}^{2}}} \approx 0$, the usual result
        (e.g.~\cite{drazin,sutherland1}). Note that for an upwards propagating
        IGW $c_{g, z} > 0$, $k_z < 0, k_x > 0$.

    \item The time-averaged total $x$-momentum flux $F$ carried in the $\bm{z}$
        direction can be computed. Since the linear solution consists of terms
        separable as $f(x, z, t) = f(z)\cos\p{k_{x}x - \omega t}$, $x$
        averaging and time averaging are equivalent. $x$ averaging is preferred
        both since periodic boundary conditions pick a natural averaging length
        and since it does not introduce extra temporal smoothing. We may then
        write
        \begin{align}
            F &\equiv \ev{\rho u_{x}' u_{z}'}_x \equiv
                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_{x}'u_{z}'\;\mathrm{d}x,
                    \label{eq:F_def},\\
                &\approx -\frac{A^2}{2}\overline{\rho}_0\frac{k_{z}}{k_{x}},
                    \label{eq:S_lin}
        \end{align}
        where $\ev{\dots}_x$ denotes $x$-averaging. Note that indeed for an
        upwards propagating IGW $k_z < 0, k_x > 0$, we have $F > 0$.
\end{itemize}

@@

A purely horizontal shear flow $\overline{U}(z) \bm{x}$ can be seen in
\autoref{se:nl_orig} to have the effect of modifying time derivatives
$\partial_t$ to their frequency in the comoving frame of the fluid $\partial_t -
\overline{U}(z)\partial_x$. If anywhere $\overline{U}(z) = \overline{U}_c$ where
\begin{equation}
    \overline{U}_c \equiv \omega / k_x,\label{eq:u_c}
\end{equation}
then the horizontal wavenumber of the linear wave in the fluid frame vanishes
and critical behavior is observed. In a linear theory, both inviscid
\citep{booker_bretherton} and viscous \citep{hazel}, the incident wave has
amplitude reflection and transmission coefficients
\begin{align}
    \mathrm{Ri} &\equiv \at{\frac{N^2}{\p{\pd{\overline{U}}{z}}^2}}_{z_c},
        \label{eq:ri_def}\\
    \mathcal{R}_A &= e^{-2\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}, &
    \mathcal{T}_A &= e^{-\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}},
        \label{eq:crit_coeffs}
\end{align}
where we have defined local Richardson number Ri at the critical layer $z_c:
\overline{U}(z_c) = \frac{\omega}{k_{x}}$. In the $\mathrm{Ri} \gg 1$ limit,
$\mathcal{R}, \mathcal{T} \ll 1$ and the incident wave is completely absorbed to
good approximation.

@@

In a realistic WD, IGW are excited very deep in the WD and propagate a
substantial number of density scale heights before reaching the breaking zone.
This propagation is very nearly exactly linear until just a few scale heights
below the breaking region. In the interest of computational feasibility, we
restrct the simulation domain to the upper WD fluid, of much smaller vertical
extent than the full physical domain. To further decrease computational cost, we
restrict ourselves to 2D, plane-parallel IGW rather than full 3D stellar
g-modes. While it is well-known wave breaking is a 3D process
(\cite{klostermeyer,winters1994}), the dynamical effect of the breaking process
is likely to be similar in 2D.

@@

We identify two manifestations of non-absorptive behavior: (i), the presence of
a reflected wave with wave vector $\bm{k} = k_{x}\bm{x} - k_{z}\bm{z}$, and (ii)
the amount of horizontal momentum flux $F$ reflected/transmitted by the critical
layer. The reflected wave amplitude and reflected flux need not agree exactly if
some reflected flux is in higher-order modes, which is indeed the case in our
simulations. Both are of physical interest, however: the reflected wave
amplitude may be of physical interest in setting up standing modes in a
realistic star, while the flux transfer properties is important for accurately
tracking angular momentum transfer during synchronization.

@@

Since reflectivity depends sensitively on accurate measurements of $\hat{A}_i,
\hat{A}_r$, we remark \autoref{eq:ahat_def} ensures orthogonality between
$k_{x}\bm{x} \pm k_{z}\bm{z}$ modes. The integral for $\hat{A}_r$ is also
performed over $z \in [z_0 + 3\sigma, z_0 + 3\sigma + H]$. Since
$\hat{A}_i(t), \hat{A}_r(t)$ vary somewhat strongly over time, we perform time
averaging over interval approximately $8\pi/\omega$, denoted by angle brackets.
We can then define the amplitude reflectivity
\begin{equation}
    \mathcal{R}_A(t) \equiv \frac{\ev{\hat{A}_r}(t)}{\ev{\hat{A}_i}(t)}
        .\label{eq:Ra_def}
\end{equation}
% 7/N per timestep, 13 timestep smoothing kernel, 24/N period

@@

Finally, we consider the impact of magnetic winding. In \citep{bukart}, magnetic
winding is used to enforce solid body rotation on the grounds that $t_A \gg
t_{gw}$, where
\begin{equation}
    t_A = \int\limits_0^R \frac{\sqrt{4\pi \rho}}{B_0}\;\mathrm{d}r
        \sim 10^2\;\mathrm{yr}\p{\frac{10^3\;\mathrm{G}}{B}}.
\end{equation}
the Alfv\'en wave crossing time (evaluated for a CO WD in \citep{fullerIV})
measures the magnetic coupling time and $t_{gw}$ measures the gravitational wave
inspiral timescale. Before solid body rotation is attained, another relevant
timescale is the synchronization timescale $t_s$. For a tidal torque $\tau$ and
tidal forcing frequency $\sigma = m\p{\Omega - \Omega_{spin}}$, we note that
angular momentum transfer is $\pd{M_{sync}}{t} \sigma R^2 = \tau$, where
$M_{sync}$ denotes the mass of the WD that has synchronized. Thus, the
synchronization timescale is
\begin{align}
    t_{sync} &\sim \frac{M_{sync}\sigma R^2}{\tau},\nonumber\\
        &\sim 2 \times 10^5\;\mathrm{yr}
            \p{\frac{M}{M_{\odot}}}
            \p{\frac{\sigma}{2\pi / (1\;\mathrm{hr})}}
            \p{\frac{R}{R_{\oplus}}}^2
            \p{\frac{10^{-14} GM_{\odot}^2/R_{\oplus}}{\tau}}.
\end{align}
A representative $\tau$ has been taken from \citep{bukart}. However, since only
the outer $\sim 10^{-4}M_{\odot}$ need be heated for the thermodynamically
interesting effects studied in \citep{fullerIV} and \citep{tidal_novae}, it
seems that magnetic winding cannot absolutely rule out energetic outbursts such
as tidal novae resulting from strong shear flows.

@@

Numerically, the nonlinear $\frac{\bm{\nabla}P}{\rho}$ term is problematic: we
desire a system where the fluid fields are not divided by one another. We
introduce $\varpi = \frac{P}{\rho}$ instead, then mandate $\overline{\rho},
\overline{\varpi}$ background fields satisfy hydrostatic equilibrium
$\bm{\nabla}\overline{\varpi} + \overline{\varpi} \bm{\nabla}\overline{\rho} +
g\bm{z} = 0$. Taking isothermal stratification, we find $\overline{\varpi} = gH$. We
further change variables to $\Upsilon = \ln \rho - \ln \overline{\rho}$ and
$\varpi' = \varpi - \overline{\varpi}$ deviations from the background state to obtain a
system of equations at most quadratic in fluid fields:
\begin{subequations}\label{se:nl_var}
    \begin{align}
        \bm{\nabla} \cdot \bm{u}' &= 0,\\
        \pd{\Upsilon}{t} + \p{\bm{u}' \cdot \bm{\nabla}} \Upsilon
            - \frac{u_z}{H} &= 0,\\
        \pd{u_{x}}{t} + \p{\bm{u}' \cdot \bm{\nabla}}u_{x}
            + \pd{\varpi'}{x} + gH\pd{\Upsilon}{x}
            + \varpi' \pd{\Upsilon}{x} &= 0,\\
        \pd{u_z}{t} + \p{\bm{u}' \cdot \bm{\nabla}}u_z
            + \pd{\varpi'}{z} + gH\pd{\Upsilon}{z}
            + \varpi' \pd{\Upsilon}{z} - \frac{\varpi'}{H} &= 0.
    \end{align}
\end{subequations}
It bears noting that these equations are exactly equivalent to the original
Euler equations and hence conserve horizontal momentum.

>>>>

    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\left\langle#1\right\rangle}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\uv}[1]{\hat{\boldsymbol{\mathbf{#1}}}}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator*{\med}{med}

@@

Compact white dwarf (WD) binary systems, with orbital periods in the range of
minutes to hours, are important for a range of astrophysical problems. They are
the most important sources of gravitational waves (GWs) for the Laser
Interferometric Space Antenna (LISA) \citep{lisa}. They are also thought to
produce interesting optical transients such as underluminous supernovae
\citep{underlum}, Ca-rich fast transients \citep{carich}, and tidal novae
\citep{tidal_novae}. Most importantly, they have been proposed as the likely
progenitors of type Ia supernovae (e.g. \citealt{Ia0,webbink} or more recently
\citealt{Ia1,Ia2}). While presently only a few tens of compact WD binaries are
known \citep{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
expand the catalog to a few hundreds \citep{lsst_wd} \citep[results based on
\emph{Gaia}'s second data release have already begun to
appear][]{gaiaDD,gaiaDD2}, and the Large Synoptic Survey Telescope (LSST, first
light scheduled for 2020) will likely detect a few thousand more
\citep{lsst_wd}. These observations will significantly advance the understanding
of WD binaries and their evolution.

@@

The Euler equations for an incompressible, barotropic fluid in a uniform
gravitational field are
\begin{subequations}\label{se:nl_orig}
    \begin{align}
        \bm{\nabla} \cdot \bm{u} &= 0,\label{eq:nl_incomp}\\
        \md{\rho}{t} &= 0 ,\label{eq:nl_density}\\
        \md{\bm{u}}{t} + \frac{\bm{\nabla}P}{\rho} + g\uv{z} &=
            0\label{eq:nl_mom}.
    \end{align}
\end{subequations}
$\md{}{t} = \pd{}{t} + \p{\bm{u} \cdot \bm{\nabla}}$ is the Lagrangian or
material derivative, and $\bm{u}, \rho, P$ denote the velocity field, density
and pressure respectively. We denote $-g\uv{z}$ constant gravitational
acceleration. Note that at hydrostatic equilibrium $\pd{}{t} = 0$ we have
$\bm{\nabla}\overline{P} = -\overline{\rho} g\uv{z}$ and so $\overline{P} =
\overline{\rho} gH$. A vertically-stratified shear flow
$\overline{u}_x(z)\uv{x}$ is permitted at hydrostatic equilibrium, but we will
assume no background flow, so $\bm{u} = \bm{u}'$. Physically, this assumption
corresponds to a non-rotating star, or going to the corotating frame of a
rigidly rotating star.

In practice, it is convenient to introduce the coordinate $\Upsilon = \ln
\frac{\rho}{\overline{\rho}}$ \citep[e.g.][]{lecoanet_anel}. This both
identically enforces $\rho > 0$ and eliminates the stiff term
$\frac{\bm{\nabla}P}{\rho}$. We also define reduced pressure $\varpi =
\frac{P}{\rho}$. Then, we may rewrite the second two equations of
\autoref{se:nl_orig} as
\begin{subequations}\label{se:nl_upsilon}
    \begin{align}
        \md{\Upsilon}{t} + u'_z \pd{\ln \overline{\rho}}{z} &= 0
            ,\label{eq:nl_up_density} \\
        \md{\bm{u}'}{t} + \bm{\nabla}\varpi
            + \varpi\bm{\nabla}\Upsilon + g\uv{z} &= 0.
    \end{align}
\end{subequations}
In the new coordinates, hydrostatic equilibrium corresponds to $\Upsilon = 0,
\overline{\varpi} = gH$.

@@

In the small perturbation limit, where flow velocities are small compared to the
characteristic space and time scales $\pd{}{t} \gg \bm{u}' \cdot
\bm{\nabla}$, we may linearize \autoref{se:nl_upsilon}. The solution in this
linear regime is given by \citep{drazin,sutherland0}:
\begin{equation}
    u'_z\p{x, z, t} = Ae^{z/2H}\cos\p{\p{k_{x}x + k_{z}z - \omega t}},
        \label{eq:k0z_sign}
\end{equation}
where $A$ is the amplitude and $\omega$ satisfies dispersion relation
\begin{equation}
    \omega^2 = \frac{N^2k_{x}^2}{k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}}.
        \label{eq:disp_rel}
\end{equation}
Our equations are valid in the limit of large sound speed $c_s \to \infty$, in
which the \emph{Brunt-V\"ais\"al\"a frequency}
\begin{equation}
    N^2 \equiv g^2\p{\rd{\rho}{P} - \frac{1}{c_s^2}} = \frac{g}{H},
\end{equation}
is constant. Other dynamical quantities are simply related to $u'_z$ \citep[see
e.g.][]{sutherland0}.

In the short-wavelength/WKB limit $\abs{k_{z}H} \gg 1$, the solution exhibits
the following characteristics:
\begin{itemize}
    \item The amplitude of the wave grows with $z$ as $e^{z/2H}$. Thus, the
        linear approximation is always violated for sufficiently large $z$.

    \item The phase and group velocities are respectively:
        \begin{align}
            \bm{c}_{p} &=
                \p{k_{x}\uv{x} + k_{z}\uv{z}}\frac{\omega}
                {k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}},\\
            \bm{c}_{g} &= N\frac{\p{k_{z}^2 + \frac{1}{4H^2}}\uv{x}
                - \p{k_{x}k_{z}\uv{z}}}
                {\p{k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}}^{3/2}}.\label{eq:vg}
        \end{align}
        We note $\bm{c}_{p} \cdot \bm{c}_g =
        \mathcal{O}\p{\frac{1}{\p{k_{z}H}^{2}}} \approx 0$, the usual result
        \citep[e.g.][]{drazin,sutherland1}. Note that for an upwards
        propagating IGW $c_{g, z} > 0$, $k_z < 0, k_x > 0$.

    \item The time-averaged total $x$-momentum flux $F$ carried in the $\uv{z}$
        direction can be computed. Since the linear solution consists of terms
        separable as $f(x, z, t) = f(z)\cos\p{k_{x}x - \omega t}$, $x$
        averaging and time averaging are equivalent. $x$ averaging is preferred
        both since periodic boundary conditions pick a natural averaging length
        and since it does not introduce extra temporal smoothing. We may then
        write
        \begin{align}
            F &\equiv \ev{\rho u_{x}' u_{z}'}_x \equiv
                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_{x}'u_{z}'\;\mathrm{d}x,
                    \label{eq:F_def},\\
                &\approx -\frac{A^2}{2}\overline{\rho}_0\frac{k_{z}}{k_{x}},
                    \label{eq:S_lin}
        \end{align}
        where $\ev{\dots}_x$ denotes $x$-averaging. Note that indeed for an
        upwards propagating IGW $k_z < 0, k_x > 0$, we have $F > 0$.
\end{itemize}

@@

A purely horizontal shear flow $\overline{U}(z) \uv{x}$ can be seen in
\autoref{se:nl_orig} to have the effect of modifying time derivatives
$\partial_t$ to their frequency in the comoving frame of the fluid $\partial_t -
\overline{U}(z)\partial_x$. If anywhere $\overline{U}(z) = \overline{U}_c$ where
\begin{equation}
    \overline{U}_c \equiv \omega / k_x,\label{eq:u_c}
\end{equation}
then the horizontal wavenumber of the linear wave in the fluid frame vanishes
and critical behavior is observed. In a linear theory, both inviscid
\citep{booker_bretherton} and viscous \citep{hazel}, the incident wave has
amplitude reflection and transmission coefficients
\begin{align}
    \mathrm{Ri} &\equiv \at{\frac{N^2}{\p{\pd{\overline{U}}{z}}^2}}_{z_c},
        \label{eq:ri_def}\\
    \mathcal{R}_A &= e^{-2\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}, &
    \mathcal{T}_A &= e^{-\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}},
        \label{eq:crit_coeffs}
\end{align}
where we have defined local Richardson number Ri at the critical layer $z_c:
\overline{U}(z_c) = \frac{\omega}{k_{x}}$. In the $\mathrm{Ri} \gg 1$ limit,
$\mathcal{R}, \mathcal{T} \ll 1$ and the incident wave is completely absorbed to
good approximation.

@@

In a realistic WD, IGW are excited very deep in the WD and propagate a
substantial number of density scale heights before reaching the breaking zone.
This propagation is very nearly exactly linear until just a few scale heights
below the breaking region. In the interest of computational feasibility, we
restrct the simulation domain to the upper WD fluid, of much smaller vertical
extent than the full physical domain. To further decrease computational cost, we
restrict ourselves to 2D, plane-parallel IGW rather than full 3D stellar
g-modes. While it is well-known wave breaking is a 3D process
\citep{klostermeyer,winters1994}, the dynamical effect of the breaking process
is likely to be similar in 2D.

@@

We identify two manifestations of non-absorptive behavior: (i), the presence of
a reflected wave with wave vector $\bm{k} = k_{x}\uv{x} - k_{z}\uv{z}$, and (ii)
the amount of horizontal momentum flux $F$ reflected/transmitted by the critical
layer. The reflected wave amplitude and reflected flux need not agree exactly if
some reflected flux is in higher-order modes, which is indeed the case in our
simulations. Both are of physical interest, however: the reflected wave
amplitude may be of physical interest in setting up standing modes in a
realistic star, while the flux transfer properties is important for accurately
tracking angular momentum transfer during synchronization.

@@

Since reflectivity depends sensitively on accurate measurements of $\hat{A}_i,
\hat{A}_r$, we remark \autoref{eq:ahat_def} ensures orthogonality between
$k_{x}\uv{x} \pm k_{z}\uv{z}$ modes. The integral for $\hat{A}_r$ is also
performed over $z \in [z_0 + 3\sigma, z_0 + 3\sigma + H]$. Since
$\hat{A}_i(t), \hat{A}_r(t)$ vary somewhat strongly over time, we perform time
averaging over interval approximately $8\pi/\omega$, denoted by angle brackets.
We can then define the amplitude reflectivity
\begin{equation}
    \mathcal{R}_A(t) \equiv \frac{\ev{\hat{A}_r}(t)}{\ev{\hat{A}_i}(t)}
        .\label{eq:Ra_def}
\end{equation}
% 7/N per timestep, 13 timestep smoothing kernel, 24/N period

@@

Finally, we consider the impact of magnetic winding. In \citep{bukart}, magnetic
winding is used to enforce solid body rotation on the grounds that $t_A \gg
t_{gw}$, where
\begin{equation}
    t_A = \int\limits_0^R \frac{\sqrt{4\pi \rho}}{B_0}\;\mathrm{d}r
        \sim 10^2\;\mathrm{yr}\p{\frac{10^3\;\mathrm{G}}{B}}.
\end{equation}
the Alfv\'en wave crossing time (evaluated for a CO WD in \citealt{fullerIV})
measures the magnetic coupling time and $t_{gw}$ measures the gravitational wave
inspiral timescale. Before solid body rotation is attained, another relevant
timescale is the synchronization timescale $t_s$. For a tidal torque $\tau$ and
tidal forcing frequency $\sigma = m\p{\Omega - \Omega_{spin}}$, we note that
angular momentum transfer is $\pd{M_{sync}}{t} \sigma R^2 = \tau$, where
$M_{sync}$ denotes the mass of the WD that has synchronized. Thus, the
synchronization timescale is
\begin{align}
    t_{sync} &\sim \frac{M_{sync}\sigma R^2}{\tau},\nonumber\\
        &\sim 2 \times 10^5\;\mathrm{yr}
            \p{\frac{M}{M_{\odot}}}
            \p{\frac{\sigma}{2\pi / (1\;\mathrm{hr})}}
            \p{\frac{R}{R_{\oplus}}}^2
            \p{\frac{10^{-14} GM_{\odot}^2/R_{\oplus}}{\tau}}.
\end{align}
A representative $\tau$ has been taken from \citep{bukart}. However, since only
the outer $\sim 10^{-4}M_{\odot}$ need be heated for the thermodynamically
interesting effects studied in \citep{fullerIV} and \citep{tidal_novae}, it
seems that magnetic winding cannot absolutely rule out energetic outbursts such
as tidal novae resulting from strong shear flows.

@@

Numerically, the nonlinear $\frac{\bm{\nabla}P}{\rho}$ term is problematic: we
desire a system where the fluid fields are not divided by one another. We
introduce $\varpi = \frac{P}{\rho}$ instead, then mandate $\overline{\rho},
\overline{\varpi}$ background fields satisfy hydrostatic equilibrium
$\bm{\nabla}\overline{\varpi} + \overline{\varpi} \bm{\nabla}\overline{\rho} +
g\uv{z} = 0$. Taking isothermal stratification, we find $\overline{\varpi} = gH$. We
further change variables to $\Upsilon = \ln \rho - \ln \overline{\rho}$ and
$\varpi' = \varpi - \overline{\varpi}$ deviations from the background state to obtain a
system of equations at most quadratic in fluid fields:
\begin{subequations}\label{se:nl_var}
    \begin{align}
        \bm{\nabla} \cdot \bm{u}' &= 0,\\
        \pd{\Upsilon}{t} + \p{\bm{u}' \cdot \bm{\nabla}} \Upsilon
            - \frac{u_z}{H} &= 0,\\
        \pd{u_{x}}{t} + \p{\bm{u}' \cdot \bm{\nabla}}u_{x}
            + \pd{\varpi'}{x} + gH\pd{\Upsilon}{x}
            + \varpi' \pd{\Upsilon}{x} &= 0,\\
        \pd{u_z}{t} + \p{\bm{u}' \cdot \bm{\nabla}}u_z
            + \pd{\varpi'}{z} + gH\pd{\Upsilon}{z}
            + \varpi' \pd{\Upsilon}{z} - \frac{\varpi'}{H} &= 0.
    \end{align}
\end{subequations}
It bears noting that these equations are exactly equivalent to the original
Euler equations and hence conserve horizontal momentum.
